Discrete Quantum Theories and Computing

### 2018-11-21 dissertation.tex
dissertation.lyx "Chapter 2 Conventional Quantum Theory" Updating...

The part of conventional quantum theory (CQT) used by quantum circuit
model is described by the following:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:-orthonormal-basis}$D$ orthonormal basis vectors for
a Hilbert space of dimension~$D$, 
\item \label{enu:probability-amplitude}the normalized $D$ complex probability
amplitude coefficients describing the contribution of each basis vector, 
\item \label{enu:unitary-matrix}a set of probability-conserving unitary
matrix operators that suffice to describe all required state transformations
of a quantum circuit, 
\item \label{enu:measurement-framework}and a measurement framework. 
\end{enumerate}
In Sec.~\ref{CQC1qubitBloch.sec}, we focus on the discrete geometric
issues raised by the properties \ref{enu:-orthonormal-basis} and
\ref{enu:probability-amplitude} given above for CQT. In Sec.~\ref{sec:fuzzy},
we introduce the important issues of \ref{enu:measurement-framework}
and the foundations of quantum probability space. The observables
and expectation values defined in Sec.~\ref{sec:fuzzy} will be used
to understand the geometry of entangled states in Sec.~\ref{sec:The-geometry-of}.
The property~\ref{enu:unitary-matrix} directly related to quantum
circuit will be introduced later in Chapter~\ref{chap:DQT=000026DQC}
when we dive into quantum algorithms.

@@

There are many things that are assumed in CQT, such as the absence
of zero norm states for non-zero vectors, and the decomposition of
complex amplitudes into a pair of ordinary real numbers. One also
typically assumes the existence of a $D$-dimensional Hilbert space
with an orthonormal basis, allowing us to write \textit{pure} states
in general as Hilbert space vectors with an Hermitian inner product:
\begin{eqnarray}
\ket{\Psi} & = & \sum_{i=0}^{D-1}\alpha_{i}\ket{i}\,.
\end{eqnarray}
Here $\alpha_{i}\in\mathComplex$ are complex probability amplitudes,
$\VVec{\alpha}\in\mathComplex^{D}$, and the $\{\ket{i}\}$ is an
orthonormal basis of states obeying $\braket{i}{k}=\delta_{ik}$.

The meaning of this is that any state $\ket{\Phi}=\sum_{i=0}^{D-1}\beta_{i}\ket{i}$
can be projected onto another state $\ket{\Psi}$ by writing 
\begin{eqnarray}
\braket{\Phi}{\Psi} & = & \sum_{i=0}^{D-1}\beta_{i}^{*}\alpha_{i}\,,
\end{eqnarray}
thus quantifying the proximity of the two states. (Here $^{*}$ denotes
complex conjugation.) This is one of many properties we take for granted
in continuum quantum mechanics that challenge us in defining a discrete
quantum geometry. To facilitate the transition to DQT carried out
in later sections, we concern ourselves first with the properties
of the simplest possible abstract state object in CQT, the single
qubit state.

@@

We can reduce 3 degrees of freedom in \Eq{A5.eq} to 2 degrees of
freedom by effectively removing $\rme^{\rmi\theta}$ (“fibering
out by the circle $\Sphere{1}$”). The standard form of this maps
(“the Hopf fibration”) is 
\begin{eqnarray}
X & = & 2\,\mathrm{Re}\ \alpha_{0}\alpha_{1}^{*}=2x_{0}x_{1}+2y_{0}y_{1}\,,\nonumber \\
Y & = & 2\,\mathrm{Im}\ \alpha_{0}\alpha_{1}^{*}=2x_{1}y_{0}-2x_{0}y_{1}\,,\label{eq:HopfFibration}\\
Z & = & \left|\alpha_{0}\right|^{2}-\left|\alpha_{1}\right|^{2}={x_{0}}^{2}+{y_{0}}^{2}-{x_{1}}^{2}-{y_{1}}^{2}\,.\nonumber 
\end{eqnarray}
By denoting the three-dimensional vector~$\left(X,Y,Z\right)$ as
$\Hat{a}$, \Eq{A5.eq} implies these transformed coordinates obeying
\begin{equation}
\left\Vert \Hat{a}\right\Vert ^{2}=X^{2}+Y^{2}+Z^{2}=\left(\left|\alpha_{0}\right|^{2}+\left|\alpha_{1}\right|^{2}\right)^{2}=1\label{A7.eq}
\end{equation}
and therefore have only two remaining degrees of freedom describing
all possible distinct one-qubit quantum states. In Fig.~\ref{hopfFibers.fig},
we illustrate schematically the family of circles \textit{each one
of which is collapsed to a point}~$\left(\phi,\psi\right)$ on the
surface $X^{2}+Y^{2}+Z^{2}=1$ by the Hopf map.

@@

@@

\begin{thm}[Gleason's theorem]\label{cor:Gleason's}In a Hilbert
space $\Hilb$ of dimension $D\geq3$, given a quantum probability
measure $\mu:\events\rightarrow\left[0,1\right]$, there exists a
unique mixed state~$\rho$ such that $\mu=\muB_{\rho}$.\end{thm}

>>>>

The part of conventional quantum theory (CQT) used by quantum circuit
model is described by the following:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:-orthonormal-basis}$D$ orthonormal basis vectors for
a Hilbert space of dimension~$D$, 
\item \label{enu:probability-amplitude}$D$ complex probability amplitude
coefficients describing the contribution of each basis vector, 
\item \label{enu:unitary-matrix}a set of probability-conserving unitary
matrix operators that suffice to describe all required state transformations
of a quantum circuit, 
\item \label{enu:measurement-framework}and a measurement framework. 
\end{enumerate}
In Sec.~\ref{CQC1qubitBloch.sec}, we focus on the geometric issues
raised by the properties \ref{enu:-orthonormal-basis} and \ref{enu:probability-amplitude}
given above for CQT. In Sec.~\ref{sec:fuzzy}, we introduce the important
issues of \ref{enu:measurement-framework} and the foundations of
quantum probability space. The observables and expectation values
defined in Sec.~\ref{sec:fuzzy} will be used to understand the geometry
of entangled states in Sec.~\ref{sec:The-geometry-of}. The property~\ref{enu:unitary-matrix}
directly related to quantum circuit will be introduced later in Chapter~\ref{chap:DQT=000026DQC}
when we dive into quantum algorithms.

@@

There are many things that are assumed in CQT, such as the absence
of zero norm states for non-zero vectors, and the decomposition of
complex amplitudes into a pair of ordinary real numbers. One also
typically assumes the existence of a $D$-dimensional Hilbert space
with an orthonormal basis, allowing us to write \textit{pure} states
in general as Hilbert space vectors with an Hermitian inner product:
\begin{equation}
\ket{\Psi}=\sum_{i=0}^{D-1}\alpha_{i}\ket{i}\,.
\end{equation}
Here $\alpha_{i}\in\mathComplex$ are complex probability amplitudes,
$\VVec{\alpha}\in\mathComplex^{D}$, and the $\{\ket{i}\}$ is an
orthonormal basis of states obeying $\braket{i}{k}=\delta_{ik}$.

The meaning of this is that any state $\ket{\Phi}=\sum_{i=0}^{D-1}\beta_{i}\ket{i}$
can be projected onto another state $\ket{\Psi}$ by writing 
\begin{equation}
\braket{\Phi}{\Psi}=\sum_{i=0}^{D-1}\beta_{i}^{*}\alpha_{i}\,,
\end{equation}
thus quantifying the proximity of the two states. (Here $^{*}$ denotes
complex conjugation.) This is one of many properties we take for granted
in continuum quantum mechanics that challenge us in defining a discrete
quantum geometry. To facilitate the transition to DQT carried out
in later sections, we concern ourselves first with the properties
of the simplest possible abstract state object in CQT, the single
qubit state.

@@

We can reduce 3 degrees of freedom in \Eq{A5.eq} to 2 degrees of
freedom by effectively removing $\rme^{\rmi\theta}$ (“fibering
out by the circle $\Sphere{1}$”). The standard form of this maps
(“the Hopf fibration”) is 
\begin{eqnarray}
 &  & X=2\,\mathrm{Re}\ \alpha_{0}\alpha_{1}^{*}=2x_{0}x_{1}+2y_{0}y_{1}\,,\nonumber \\
 &  & Y=2\,\mathrm{Im}\ \alpha_{0}\alpha_{1}^{*}=2x_{1}y_{0}-2x_{0}y_{1}\,,\label{eq:HopfFibration}\\
 &  & Z=\left|\alpha_{0}\right|^{2}-\left|\alpha_{1}\right|^{2}={x_{0}}^{2}+{y_{0}}^{2}-{x_{1}}^{2}-{y_{1}}^{2}\,.\nonumber 
\end{eqnarray}
By denoting the three-dimensional vector~$\left(X,Y,Z\right)$ as
$\Hat{a}$, \Eq{A5.eq} implies these transformed coordinates obeying
\begin{equation}
\left\Vert \Hat{a}\right\Vert ^{2}=X^{2}+Y^{2}+Z^{2}=\left(\left|\alpha_{0}\right|^{2}+\left|\alpha_{1}\right|^{2}\right)^{2}=1\label{A7.eq}
\end{equation}
and therefore have only two remaining degrees of freedom describing
all possible distinct one-qubit quantum states. In Fig.~\ref{hopfFibers.fig},
we illustrate schematically the family of circles \textit{each one
of which is collapsed to a point}~$\left(\phi,\psi\right)$ on the
surface $X^{2}+Y^{2}+Z^{2}=1$ by the Hopf map.

@@

After introducing quantum probability measures, we might follow the
convention to introduce the Born rule which is the only way to construct
quantum probability measures in CQT. However, since the variants of
quantum probability measures in Sec.~\ref{sec:Toward-IVPM} and Chapter~\ref{chap:QIVPM}
could not be constructed by the Born rule easily, it is instructive
to search for a Born rule here, and we will actually find the unique
one.

Although quantum probability measures constructed by gluing classical
probability measures in the previous section looks complex, it could
actually be constructed by operators according to Gleason's theorem.

\begin{thm}[Gleason's theorem]\label{cor:Gleason's}In a Hilbert
space $\Hilb$ of dimension $D\geq3$, given a quantum probability
measure $\mu:\events\rightarrow\left[0,1\right]$, there exists a
unique mixed state~$\rho$ such that $\mu\left(P\right)=\Tr\left(\rho P\right)$
for any $D$-dimensional projector~$P$.\end{thm}

\noindent In another word, when sending a particle to a Stern-Gerlach
type experiment, we will get a quantum probability measure of the
form $\Tr\left(\rho P\right)$ for some operator~$\rho$.

Recall in Sec.~\ref{CQC1qubitBloch.sec}, the status of our system
can be characterized by

@@
